\newcommand{\etalchar}[1]{$^{#1}$}
\begin{thebibliography}{KMS{\etalchar{+}}11b}

\bibitem[Aho96]{ahonen}
Helena~Ahonen.
\newblock {\em Generating Grammars for Structured Documents Using Grammatical Inference Methods}.
\newblock PhD thesis, Department of Computer Science, University of Helsinki, Series of Publications A, Report A-1996-4, 1996.

\bibitem[BDF{\etalchar{+}}00]{fajt-12}
Peter Buneman, Susan Davidson, Wenfei Fan, Carmem Hara, and Wang chiew Tan.
\newblock Reasoning About Keys for XML.
\newblock In {\em Information Systems}, page 2003, 2000.

\bibitem[BDF{\etalchar{+}}01]{keX}
Peter Buneman, Susan Davidson, Wenfei Fan, Carmem Hara, and~Wang-Chiew Tan.
\newblock Keys for XML.
\newblock In {\em Proceedings of the 10th international conference on World Wide Web}, WWW '01, pages 201--210, New York, NY, USA, 2001. ACM.

\bibitem[BLNT07]{fajt-51}
Jana Bauckmann, Ulf Leser, Felix Naumann, and Véronique Tietz.
\newblock Efficiently Detecting Inclusion Dependencies.
\newblock In {\em In Int. Conf. on Data Engineering (ICDE 07}. Poster, 2007.

\bibitem[BM03]{fidax}
Denilson Barbosa and Alberto Mendelzon.
\newblock Finding ID Attributes in XML Documents.
\newblock In Zohra Bellahsene, Akmal Chaudhri, Erhard Rahm, Michael Rys, and Rainer Unland, editors, {\em Database and~XML Technologies}, volume 2824 of~{\em Lecture Notes in Computer Science}, pages 180--194. Springer Berlin / Heidelberg, 2003.
\newblock 10.1007/978-3-540-39429-7\_12.

\bibitem[BNV08]{bex}
Geert~Jan Bex, Frank Neven, and Stijn Vansummeren.
\newblock Schemascope: a System for Inferring and Cleaning XML Schemas.
\newblock In {\em Proceedings of the 2008 ACM SIGMOD international conference on Management of data}, SIGMOD '08, pages 1259--1262, New York, NY, USA, 2008. ACM.

\bibitem[BPM{\etalchar{+}}08]{Bray:08:EML}
Tim Bray, Jean Paoli, Eve Maler, Fran\c{c}ois Yergeau, and C.~M. Sperberg-McQueen.
\newblock Extensible Markup Language ({XML}) 1.0 (Fifth Edition).
\newblock {W3C} recommendation, W3C, November 2008.
\newblock \url{http://www.w3.org/TR/2008/REC-xml-20081126/}.

\bibitem[Dan98]{dantzig1998linear}
G.B. Dantzig.
\newblock {\em Linear Programming and Extensions}.
\newblock Landmarks in~Physics and~Mathematics. Princeton University Press, 1998.

\bibitem[DC10]{heu-lecture}
G.~Di~Caro.
\newblock Heuristics Lecture Notes.
\newblock 2010.

\bibitem[DS04]{dorigo2004ant}
M.~Dorigo and T.~St{\"u}tzle.
\newblock {\em Ant Colony Optimization}.
\newblock Bradford Books. MIT Press, 2004.

\bibitem[Faj10]{fajt}
Stanislav Fajt.
\newblock Mining XML Integrity Constraints.
\newblock Master's thesis, Charles University in Prague, 2010.

\bibitem[FGK09]{Fomin:2009:MCA:1552285.1552286}
Fedor~V. Fomin, Fabrizio Grandoni, and Dieter Kratsch.
\newblock A Measure \& Conquer Approach for the Analysis of Exact Algorithms.
\newblock {\em J. ACM}, 56:25:1--25:32, August 2009.

\bibitem[GL97]{Glover:1997:TS:549765}
Fred Glover and Manuel Laguna.
\newblock {\em Tabu Search}.
\newblock Kluwer Academic Publishers, Norwell, MA, USA, 1997.

\bibitem[glp]{glpk}
GNU Linear Programming Kit.
\newblock \url{http://www.gnu.org/s/glpk/}.

\bibitem[gnu]{gnuplot}
Gnuplot, an Interactive Plotting Program.\\
\newblock \url{http://www.gnuplot.info/}.

\bibitem[Gol89]{goldberg1989genetic}
D.E. Goldberg.
\newblock {\em Genetic Algorithms in Search, Optimization, and Machine Learning}.
\newblock Artificial Intelligence. Addison-Wesley Pub. Co., 1989.

\bibitem[gra]{graphviz}
Graph Visualization Software.
\newblock \url{http://www.graphviz.org/}.

\bibitem[GZ02]{fajt-39}
G\"{o}sta Grahne and Jianfei Zhu.
\newblock Discovering Approximate Keys in XML Data.
\newblock In {\em Proceedings of the eleventh international conference on Information and knowledge management}, CIKM '02, pages 453--460, New York, NY, USA, 2002. ACM.

\bibitem[JR86]{JM1986425}
J.M and Robson.
\newblock Algorithms for Maximum Independent Sets.
\newblock {\em Journal of Algorithms}, 7(3):425 -- 440, 1986.

\bibitem[JUF10]{fajt-53}
B.~Jana, L.~Ulf, and N.~Felix.
\newblock {\em Efficient and Exact Computation of Inclusion Dependencies for Data Integration}.
\newblock Technische Berichte des Hasso-Plattner-Instituts f{\"u}r Softwaresystemtechnik an der Universit{\"a}t Potsdam. Universit{\"a}tsverlag Potsdam, 2010.

\bibitem[Kel]{gnuplot-manual}
Williams Kelley.
\newblock Gnuplot Manual.
\newblock \url{http://www.gnuplot.info/}.

\bibitem[Kle11]{anti}
Michal Klempa.
\newblock Optimization and Refinement of XML Schema Inference Approaches.
\newblock Master's thesis, Charles University in Prague, 2011.

\bibitem[KMS{\etalchar{+}}11a]{archdoc}
Michal Klempa, M\'{a}rio Mikula, Robert Smetana, Michal \v{S}virec, and Matej Vit\'{a}sek.
\newblock jInfer Architecture.\\
\newblock \url{http://jinfer.sourceforge.net/modules/architecture.pdf}, 2011.

\bibitem[KMS{\etalchar{+}}11b]{basiciggdoc}
Michal Klempa, M\'{a}rio Mikula, Robert Smetana, Michal \v{S}virec, and Matej Vit\'{a}sek.
\newblock jInfer BasicIGG Module Description.\\
\newblock \url{http://jinfer.sourceforge.net/modules/basicigg.pdf}, 2011.

\bibitem[KMS{\etalchar{+}}11c]{jinferweb}
Michal Klempa, M\'{a}rio Mikula, Robert Smetana, Michal \v{S}virec, and Matej Vit\'{a}sek.
\newblock jInfer: Java Framework for XML Schema Inference.\\
\newblock \url{http://jinfer.sourceforge.net}, 2011.

\bibitem[LD60]{land60a}
A.~H. Land and A.~G Doig.
\newblock An Automatic Method of Solving Discrete Programming Problems.
\newblock {\em Econometrica}, 28(3):497--520, 1960.

\bibitem[Mak]{mathprog}
Andrew Makhorin.
\newblock GNU MathProg.
\newblock \url{http://lpsolve.sourceforge.net/5.5/MathProg.htm}.

\bibitem[McK66]{birthday}
E.~H. McKinney.
\newblock Generalized Birthday Problem.
\newblock {\em The American Mathematical Monthly}, 73(4):pp. 385--387, April 1966.

\bibitem[Nor]{extendedcfg}
Theodore Norvell.
\newblock A Short Introduction to Regular Expressions and Context-Free Grammars.\\
\newblock \url{http://www.engr.mun.ca/~theo/Courses/fm/pub/context-free.pdf}.

\bibitem[Pas97]{Paschos:1997:SAO:254180.254190}
Vangelis~T. Paschos.
\newblock A Survey of Approximately Optimal Solutions to Some Covering and Packing Problems.
\newblock {\em ACM Comput. Surv.}, 29:171--209, June 1997.

\bibitem[SBHR06]{fajt-41}
Yannis Sismanis, Paul Brown, Peter~J. Haas, and Berthold Reinwald.
\newblock Gordian: Efficient and Scalable Discovery of Composite Keys.
\newblock In {\em In Proc. International Conference on Very Large Data Bases (VLDB}, pages 691--702, 2006.

\bibitem[Vli02]{vlist2002xml}
E.V. Vlist.
\newblock {\em XML Schema}.
\newblock O'Reilly Series. O'Reilly, 2002.

\bibitem[VMP08]{1802522}
Ond\v{r}ej Vo\v{s}ta, Irena Ml\'{y}nkov\'{a}, and Jaroslav Pokorn\'{y}.
\newblock Even an Ant Can Create an XSD.
\newblock In {\em DASFAA'08: Proceedings of the 13th international conference on Database systems for advanced applications}, pages 35--50, Berlin, Heidelberg, 2008. Springer-Verlag.

\bibitem[\v{S}11]{sviro}
Michal \v{S}virec.
\newblock Efficient Detection of XML Integrity Constraints.
\newblock Master's thesis, Charles University in Prague, 2011.

\bibitem[Vyh09]{vyhnanovska}
Julie Vyhnanovsk\'{a}.
\newblock Automatic Construction of an XML Schema for a Given Set of XML Documents.
\newblock Master's thesis, Charles University in Prague, 2009.

\end{thebibliography}