\chapter{Related Work}
\label{chapter-research}

According to the article \cite[Chapter~4]{fidax}, the problem of finding an ID set with weight more than some $K$ ($K$-\textsc{IDSet}) is in NP. Furthermore, the independent set (IS) problem can be reduced to $K$-\textsc{IDSet}, meaning $K$-\textsc{IDSet} is NP-hard and thus NP-complete. The transformation from IS problem formulation to~$K$-\textsc{IDSet} problem formulation is as follows.

\begin{quote}
Let $G = (V, E)$ be a simple connected graph with vertex set $V = \{v_1, \ldots, v_n\}$, and edge set $E = \{e_1, \ldots, e_m\}$. We define the attribute mappings as follows. Let $ \mathcal{I} = V \cup E$, and define $value(x) = x, x \in \mathcal{I}$. For each vertex $v_i \in V$, we create a mapping $m_i = \{(v_i, e_j): e_j \in E \,\text{is incident on}\, v_i \}$, and define $\tau(m_i) = v_i$; let $C = \{m_1, \ldots, m_n\}$ be set of all such mappings. It is clear that G has an independent set of size $K$ iff $C$ has an ID set of size $K$. Also, $C$ can be constructed in time polynomial on $n+m$.
\end{quote}

The article continues by proving that finding the maximum weighted IS can be reduced to the problem of finding an ID set with maximum weight (\textsc{Max-IDSet}). This again means that \textsc{Max-IDSet} is NP-complete and, furthermore, unless $\P = \NP$, \textsc{Max-IDSet} has no constant factor approximation algorithm.

The difference in transformation from maximum weighted IS to \textsc{Max-IDSet} is as follows.

\begin{quote}
[...] with the added restriction that $w(m_i) = w(v_i), v_i \in V$.
\end{quote}

Note that the transformation works in both ways: it is equivalently possible to create a maximum weighted IS instance for a given \textsc{Max-IDSet} instance.

The article further suggests a heuristic approach described in Section \ref{section-mip-fidax}, which was incorporated into the framework proposed by this work.

To the best of our knowledge, there are no other articles dealing with this problem.\\

\section{Finding XML Keys}

XML keys are a structure somewhat similar to \texttt{ID} attributes, but with a much larger expressive strength. They have been introduced in \cite{keX} and implemented in XML Schema\footnote{\url{http://www.w3.org/TR/xmlschema11-1/\#Identity-constraint\_Definition\_details}}.

Fajt in \cite{fajt} summarizes several algorithms to help find XML keys in existing data, namely \textit{Gordian}, \textit{XML Primary Keys}, \textit{SPIDER} and \textit{DBA Companion}. Except for \textit{XML Primary Keys}, they all are originally purposed to find keys in~relational databases. We will describe them shortly.

\subsubsection{\textit{Gordian}}

This algorithm from \cite{fajt-41} extracts composite primary keys (PKs) from relational databases.

\begin{quote}
The idea behind is an observation that a projection of entities corresponds to a key if each counted aggregation for a projection is equal to 1. Thus, this method searches for all possible projections of a dataset while computing aggregations on the projected part of the set of entities.
\end{quote}

This is achieved by constructing a prefix tree from the tuples in the original relation, which is then pruned and traversed depth-first to find non-key attributes from which the primary keys are inferred. This algorithm still has to~be~adapted to~search for PKs in XML data.

\subsubsection{\textit{XML Primary Keys}}

This is an algorithm from \cite{fajt-39} capable of finding simple keys and foreign keys directly in XML data. This is achieved by building a prefix tree containing all the XML nodes and then evaluating every path in it as a candidate key using metrics called \textit{support} and \textit{confidence}. To find more complex keys, the algorithm iteratively constructs candidate keys from simpler ones and evaluates them.\\

The following two algorithms deal with \textit{inclusion dependencies} (INDs), described for example in \cite{fajt-12}.

\subsubsection{\textit{SPIDER}}

The core of this algorithm from \cite{fajt-51, fajt-53} is the following.

\begin{quote}
The process consists of two steps - sets of values are sorted during the first one and then all the candidates are analyzed in parallel. The core of the method is utilizing the data structure called min-heap which synchronizes the processing of all values of all attributes.
\end{quote}

It is~possible to~use a~number of~heuristic pruning strategies to~keep the~min-heap in~a~reasonable size. This algorithm performs very well for PKs in~relational databases, however, it~still has to~be~adapted for~XML keys.

\subsubsection{\textit{DBA Companion}}

Like \textit{SPIDER}, this method from \cite{fajt-53} is able to find all the~INDs in~the~da\-ta\-base in~just one pass. However, it~uses a~different data structure (basically a~binary relation between the~attributes and~their corresponding values) and~considers data types. Composite INDs are~found using the~simple ones and~pruning the~search space. According to~the~authors of~\textit{SPIDER}, \textit{DBA Companion} is~far inferior in~performance. This algorithm has yet to~be~adapted to~search for XML keys, too.

\subsubsection{Fajt's Approach - \textit{KeyMiner}}

Fajt introduces a new algorithm based on \textit{Gordian} and \textit{SPIDER} to look for primary and foreign keys in XML data. First, relations have to be extracted from the original XML document. Then all the primary keys are found using a modified \textit{Gordian} algorithm which can find absolute as well as relative PKs. Finally, \textit{SPIDER} is used to compute the foreign keys from the PKs found in the previous step.

\subsection{Relation to \texttt{ID} Attributes}

XML keys found this or any other way can under some circumstances (when they are simple enough) be translated to an equivalent \texttt{ID} attribute definition. The process is described in \cite[Ch.\ 9, s.\ 3]{vlist2002xml}. This opens a new line of possible research: finding XML keys using an algorithm modified to look only for \textit{useful} keys and then converting them to \texttt{ID} attributes.\\ 

However, in our work we find \texttt{ID} attributes directly. And even though we can always convert them to XML keys by the process mentioned above, we are unable to find more complex keys this way.\\

\section{Maximum Weighted IS}

Maximum weigthed IS is a well researched topic with a lot of known direct or approximation algorithms, see e.g. \cite{JM1986425} or \cite{Fomin:2009:MCA:1552285.1552286}. According to for example \cite{Paschos:1997:SAO:254180.254190}, the best known approximation algorithm for weighted IS to-date achieves an approximation ratio of $3(\Delta + 2)$, where $\Delta$ is the maximum degree of a vertex in the IS graph. This article lists several algorithms similar to those we introduce in the following chapters.\\